### 2019-06-08 1-Preamble/5-Abstract/abstract.tex
Starting to read everything to hunt := notations

% ************************** Thesis Abstract *****************************
% Use `abstract' as an option in the document class to print only the titlepage and the abstract.
\begin{abstract}
\vspace*{-30pt}
%
% \TODOL{Raccourcir juste un peu ?}
% ----------------------------------------------------------------------------
\begin{small}
% Chapitre 1
Dans cette thèse de doctorat, nous étudions les réseaux sans fil et les appareils reconfigurables qui peuvent accéder à des réseaux de radio intelligente, dans des bandes non licenciées et sans supervision centrale.
Nous considérons des réseaux de l'Internet des Objets (IoT), avec l'objectif d'augmenter la durée de vie de la batterie des appareils, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil.
%  (Chapitre~\ref{chapter:1}).
% Chapitre 4
Nous proposons différents modèles de réseaux IoT, et nous montrons empiriquement, par des simulations numériques et une validation expérimentale réaliste, le gain que peuvent apporter nos méthodes, qui utilisent l'apprentissage par renforcement.
%  (Chapitre~\ref{chapitre:4}).
% Chapitre 2
Les différents problèmes d'accès au réseau sont modélisés avec des Bandits Multi-Armés (MAB), mais leur analyse est difficile à réaliser,
% , Chapitre~\ref{chapitre:2}
car il est difficile de prouver la convergence de nombreux appareils jouant à un jeu collaboratif sans communication aucune coordination, lorsque les appareils suivent tous un modèle d'activation aléatoire.
% En effet, même le nombre de périphériques actifs à chaque instant est hautement imprévisible, rendant l'environnement non pas stationnaire et évoluant rapidement.
% Chapitres 5 et 6
Le reste de ce manuscrit étudie donc deux modèles restreints, d'abord des bandits multi-joueurs dans des problèmes stationnaires, puis des bandits simple-joueurs non stationnaires.
%  (Chapitre~\ref{chapitre:5})
%  (Chapitre~\ref{chapitre:6})
% Chapitre 3
Nous détaillons également une autre contribution, la bibliothèque Python open-source SMPyBandits pour des simulations numériques de problèmes MAB, qui couvre tous les modèles étudiés et d'autres.
% (Chapitre~\ref{chapter:3})
\end{small}

>>>>

% ************************** Thesis Abstract *****************************
% Use `abstract' as an option in the document class to print only the titlepage and the abstract.
% \vspace*{-30pt}
%
\begin{abstract}
\vspace*{-30pt}
%
% \TODOL{Raccourcir juste un peu ?}
% ----------------------------------------------------------------------------
\begin{small}
% Chapitre 1
Dans cette thèse de doctorat, nous étudions les réseaux sans fil et les appareils reconfigurables qui peuvent accéder à des réseaux de radio intelligente, dans des bandes non licenciées et sans supervision centrale.
Nous considérons des réseaux de l'Internet des Objets (IoT), avec l'objectif d'augmenter la durée de vie de la batterie des appareils, en les équipant d'algorithmes d'apprentissage machine peu coûteux mais efficaces, qui leur permettent d'améliorer automatiquement l'efficacité de leurs communications sans fil.
%  (Chapitre~\ref{chapter:1}).
% Chapitre 4
Nous proposons différents modèles de réseaux IoT, et nous montrons empiriquement, par des simulations numériques et une validation expérimentale réaliste, le gain que peuvent apporter nos méthodes, qui utilisent l'apprentissage par renforcement.
%  (Chapitre~\ref{chapitre:4}).
% Chapitre 2
Les différents problèmes d'accès au réseau sont modélisés avec des Bandits Multi-Armés (MAB), mais leur analyse est difficile à réaliser,
% , Chapitre~\ref{chapitre:2}
car il est difficile de prouver la convergence de nombreux appareils jouant à un jeu collaboratif sans communication ni aucune coordination, lorsque les appareils suivent tous un modèle d'activation aléatoire.
% En effet, même le nombre de périphériques actifs à chaque instant est hautement imprévisible, rendant l'environnement non pas stationnaire et évoluant rapidement.
% Chapitres 5 et 6
Le reste de ce manuscrit étudie donc deux modèles restreints, d'abord des bandits multi-joueurs dans des problèmes stationnaires, puis des bandits simple-joueurs non stationnaires.
%  (Chapitre~\ref{chapitre:5})
%  (Chapitre~\ref{chapitre:6})
% Chapitre 3
Nous détaillons également une autre contribution, la bibliothèque Python open-source SMPyBandits pour des simulations numériques de problèmes MAB, qui couvre tous les modèles étudiés et d'autres.
% (Chapitre~\ref{chapter:3})
\end{small}

### 2019-06-08 1-Preamble/5-Abstract/resume_fr.tex
Starting to read everything to hunt := notations

@@

    \item
    Le code source des deux contributions susmentionnées utilise GNU Octave ou MATLAB et est publié en ligne, ainsi que des instructions claires pour la reproduction de nos travaux.

@@

>>>>

\TODOL{Je ne sais pas quelle longueur ni quel niveau de détail il faut viser ici. Est-ce que ces 3/4 pages suffisent ? Est-ce qu'il faut donner des maths en français pour CHAQUE CHAPITRE ? J'ai un peu la flemme d'écrire 20 pages avec des maths en français juste pour pouvoir citer tel théorème ou tel résultat.}

@@

    \item
    Le code source des deux contributions susmentionnées est publié en ligne, ainsi que des instructions pour la reproduction de nos travaux (avec GNU Octave ou MATLAB).

@@

% WARNING
\vfill{}

\textbf{Note sur le droit intellectuel.}
%
Ce document et les ressources additionnelles requises pour le compiler (notamment les fichiers \LaTeX, les morceaux de code Python, les figures etc)
sont \href{https://github.com/Naereen/phd-thesis/}{publiées publiquement},
selon les termes de la \href{https://lbesson.mit-license.org/}{\emph{licence MIT}} open-source,
en ligne sur \href{https://github.com/Naereen/phd-thesis/}{\texttt{GitHub.com/Naereen/phd-thesis/}}.

\TODOL{Open source the repository as soon as I defended my thesis!}

\begin{center}
    \textbf{Copyright 2016-2019, \copyright ~Lilian~Besson.}
\end{center}

### 2019-06-08 2-Chapters/1-Chapter/chapter1.tex
Starting to read everything to hunt := notations

\paragraph{Abstract.}
%
We start by presenting the problems studied in this thesis,
then we announce the different directions we considered and the solutions we proposed.
We give an overview of our main contributions, alongside a list of publications written in the last three years.
%
The organization of this manuscript is detailed in a graph, which highlights that Chapters~\ref{chapter:4} to \ref{chapter:6} are partly independent, but use the notations introduced on Chapter~\ref{chapter:2}.

@@

\textbf{Seminars.}
    I gave some presentations in the following events:
    SequeL team seminar at Inria Lille in September and December $2017$, and June $2019$;
    SCEE team seminar at CentraleSupélec, Rennes campus, in October $2017$, February $2018$ and June $2019$;
    as well as
    for the GDR ISIS day held in Issy-les-Moulineaux on November $2017$,
    for the brown-bag seminar at ENSAI in Bruz in January $2018$,
    for the weekly seminar at CMAP lab at École Polytechnique in October $2018$,
    for the weekly seminar of the PANAMA project-team at IRISA / Inria Rennes in May $2019$.

\textbf{Tutorial.}
    I gave a tutorial on the Julia language, at IETR seminar in Vannes in June $2018$,
    with Pierre Haessig (\texttt{pierreh.eu}), see \href{https://HAL.Inria.fr/cel-01830248}{\texttt{HAL.Inria.fr/cel-01830248}}.

>>>>

\paragraph{Abstract.}
%
We start by presenting the problems studied in this thesis,
then we announce the different directions we considered and the solutions we proposed.
We give an overview of our main contributions, alongside a list of publications written in the last three years.
%
The organization of this manuscript is detailed in a graph, which highlights that Chapters~\ref{chapter:4} to \ref{chapter:6} are partly independent, but all use the notations introduced on Chapter~\ref{chapter:2}.

@@

\textbf{Seminars.}
    I gave some presentations in the following events:
    SequeL team seminar at Inria Lille in September and December $2017$, and June $2019$;
    SCEE team seminar at CentraleSupélec, Rennes campus, in October $2017$, February $2018$ and June $2019$;
    as well as
    for the GDR ISIS day held in Issy-les-Moulineaux on November $2017$,
    for the brown-bag seminar at ENSAI in Bruz in January $2018$,
    for the weekly seminar at CMAP lab at École Polytechnique in October $2018$,
    for the weekly seminar of the PANAMA project-team at IRISA / Inria Rennes in June $2019$.

\textbf{Tutorial.}
    I gave a tutorial on the Julia language, at IETR seminar in Vannes in June $2018$,
    with Pierre Haessig (see \href{https://pierreh.eu/}{\texttt{pierreh.eu}}), see \href{https://HAL.Inria.fr/cel-01830248}{\texttt{HAL.Inria.fr/cel-01830248}}.

### 2019-06-08 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Starting to read everything to hunt := notations

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.70\linewidth]{system_model1.eps}
    \caption{In our system model, some dynamic devices (in the \textcolor{blue}{IoT network in blue}) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in \textcolor{orange}{orange}).}
    \label{fig:41:system_model1}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[scale=0.40]{protocol.eps}
    \caption{The considered time-frequency slotted protocol. Each frame is composed by a fixed duration \textcolor{blue}{uplink slot} in which the end-devices transmit their \textcolor{blue}{(uplink) packets}. If a packet is well received, the gateway replies by transmitting an \textcolor{darkgreen}{\Ack}, after the ack delay.}
    \label{fig:41:protocol}
\end{figure}

>>>>

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.70\linewidth]{system_model1.eps}
    \caption[In our system model, some dynamic devices (in the IoT network in blue) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in orange).]{In our system model, some dynamic devices (in the \textcolor{blue}{IoT network in blue}) transmit packets to a gateway and suffer from the interference generated by neighboring networks (in \textcolor{orange}{orange}).}
    \label{fig:41:system_model1}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[scale=0.40]{protocol.eps}
    \caption[The considered time-frequency slotted protocol. Each frame is composed by a fixed duration uplink slot in which the end-devices transmit their (uplink) packets. If a packet is well received, the gateway replies by transmitting an \Ack, after the ack delay.]{The considered time-frequency slotted protocol. Each frame is composed by a fixed duration \textcolor{blue}{uplink slot} in which the end-devices transmit their \textcolor{blue}{(uplink) packets}. If a packet is well received, the gateway replies by transmitting an \textcolor{darkgreen}{\Ack}, after the ack delay.}
    \label{fig:41:protocol}
\end{figure}

### 2019-06-08 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits_appendix.tex
Starting to read everything to hunt := notations

%
% System regret and three terms
%
\begin{figure}[!h]
    \centering
    % \begin{subfigure}[!h]{1.00\textwidth}
        \includegraphics[width=1.00\textwidth]{main_RegretCentralized____env4-4_2092905764868974160.pdf}
    % \end{subfigure}
    % ~
    % \begin{subfigure}[!h]{1.00\textwidth}
        \includegraphics[width=1.00\textwidth]{main_RegretCentralized____env4-4_1699475021767911583.pdf}
    % \end{subfigure}
    \caption[Multi-player regret with its three terms]{Regret with its three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3}, and lower bounds \eqref{eq:5:ourLowerBound} and \eqref{eq:5:Zhao10LowerBound} in \textbf{black}, for \Selfish-\klUCB: $M=6$ and $M=9$ players, $K=9$ arms, horizon $T=10000$ (for $1000$ runs).}
    \label{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6}
    % \vspace*{-15pt}  % XXX remove if problem
\end{figure}

>>>>

%
% System regret and three terms
%
\begin{figure}[!h]
    \centering
    % \begin{subfigure}[!h]{1.00\textwidth}
        \includegraphics[width=1.00\textwidth]{main_RegretCentralized____env4-4_2092905764868974160.pdf}
    % \end{subfigure}
    % ~
    % \begin{subfigure}[!h]{1.00\textwidth}
        \includegraphics[width=1.00\textwidth]{main_RegretCentralized____env4-4_1699475021767911583.pdf}
    % \end{subfigure}
    \caption[Illustrating the decomposition of the multi-player regret with its three terms]{Illustrating the decomposition of the multi-player regret with its three terms \ref{eq:5:term1}, \ref{eq:5:term2}, \ref{eq:5:term3}, and lower bounds \eqref{eq:5:ourLowerBound} and \eqref{eq:5:Zhao10LowerBound} in \textbf{black}, for \Selfish-\klUCB: $M=6$ and $M=9$ players, $K=9$ arms, horizon $T=10000$ (for $1000$ runs).}
    \label{fig:5:MP__M9_K9_T10000_N1000__9_algos__main_RegretCentralized____env6}
    % \vspace*{-15pt}  % XXX remove if problem
\end{figure}

### 2019-06-08 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Starting to read everything to hunt := notations

\begin{figure}[h!]  % [htbp]
    \centering
    \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_1.pdf}
    \subcaption{\textbf{Problem $1$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on only one arm at a time (\ie, $C=4$).
        The means are in $[0,1]$, and there are $C+1=5$ stationary intervals of equal lengths.
        Some changes do not modify the optimal arm (\eg, at $T=1000$ and $T=4000$) and others do.
    }
    \label{fig:6:Problem_1}
\end{figure}

@@

\begin{figure}[h!]  % [htbp]
    \centering
    \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_2.pdf}
    \subcaption{\textbf{Problem $2$}: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on all arms (\ie, $C=12$).
        The means are again in $[0,1]$, and there are also $C+1=5$ stationary intervals of equal lengths.
        % Some changes do not modify the optimal arm, but the mean of each arm does change at all change points.
    }
    \label{fig:6:Problem_2}
\end{figure}

@@

\begin{figure}[h!]  % [htbp]
    \centering
    \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_5.pdf}
    \caption{\textbf{Problem $3$}: $K=6$, $T=20000$, $C=19$ changes occur on most arms at $\Upsilon=8$ break-points.}
    \label{fig:6:Problem_5}
\end{figure}

@@

\begin{figure}[h!]  % [htbp]
    \centering
    \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_4.pdf}
    \caption{\textbf{Problem $4$}: $K=3$, $T=5000$, $C=12$ changes occur on all arms at $\Upsilon=4$ break-points.}
    % \caption{\textbf{Problem 4}: $K=3$ arms with $T=5000$, $\Upsilon=4$ changes occur on all arms at a time (\ie, $C=12$).}
    \label{fig:6:Problem_4}
\end{figure}

@@

\begin{figure}[h!]  % [htbp]
    \centering
    \includegraphics[width=0.95\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_6.pdf}
    \caption{\textbf{Pb $5$}: $K=5$, $T=100000$, $C=179$ changes occur on some arms at $\Upsilon=81$ break-points.}
    % \caption{\textbf{Problem 5}: $K=5$, $T=100000$, $\Upsilon=81$ changes occurring on some arms at every time ($C=179$).}
    \label{fig:6:Problem_6}
\end{figure}

>>>>

\begin{figure}[h!]  % [htbp]
    \centering
    \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_1.pdf}
    \caption{Problem $1$: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on only one arm at a time (\ie, $C=4$).
        The means are in $[0,1]$, and there are $C+1=5$ stationary intervals of equal lengths.
        Some changes do not modify the optimal arm (\eg, at $T=1000$ and $T=4000$) and others do.
    }
    \label{fig:6:Problem_1}
\end{figure}

@@

\begin{figure}[h!]  % [htbp]
    \centering
    \includegraphics[width=1.05\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_2.pdf}
    \caption{Problem $2$: $K=3$ arms with $T=5000$, and $\Upsilon=4$ changes occur on all arms (\ie, $C=12$).
        The means are again in $[0,1]$, and there are also $C+1=5$ stationary intervals of equal lengths.
        % Some changes do not modify the optimal arm, but the mean of each arm does change at all change points.
    }
    \label{fig:6:Problem_2}
\end{figure}

@@

\begin{figure}[h!]  % [htbp]
    \centering
    \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_5.pdf}
    \caption{Problem $3$: $K=6$, $T=20000$, $C=19$ changes occur on most arms at $\Upsilon=8$ break-points.}
    \label{fig:6:Problem_5}
\end{figure}

@@

\begin{figure}[h!]  % [htbp]
    \centering
    \includegraphics[width=0.90\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_4.pdf}
    \caption{Problem $4$: $K=3$, $T=5000$, $C=12$ changes occur on all arms at $\Upsilon=4$ break-points.}
    % \caption{\textbf{Problem 4}: $K=3$ arms with $T=5000$, $\Upsilon=4$ changes occur on all arms at a time (\ie, $C=12$).}
    \label{fig:6:Problem_4}
\end{figure}

@@

\begin{figure}[h!]  % [htbp]
    \centering
    \includegraphics[width=0.95\linewidth]{2-Chapters/6-Chapter/nonstatbandits.git/figures/6_Problems/Problem_6.pdf}
    \caption{Pb $5$: $K=5$, $T=100000$, $C=179$ changes occur on some arms at $\Upsilon=81$ break-points.}
    % \caption{\textbf{Problem 5}: $K=5$, $T=100000$, $\Upsilon=81$ changes occurring on some arms at every time ($C=179$).}
    \label{fig:6:Problem_6}
\end{figure}
